---
title: "Group WRK"
author: "Waimiri Joseph Romeo 665258 Nzambuli Daniel 665721"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Bar chart(%) patient highest Completed Education Level

```{r}
library(readxl)
library(dplyr)
library(stringr)
library(ggplot2)
library(ggpubr)

Patient_Data <- read_excel("Task2_data.xlsx")
head(Patient_Data)
```

```{r}
calculate_percentage <- function(data, column_name) {
  data <- as.data.frame(table(data[[column_name]]))
  data$Percentage = data$Freq / sum(data$Freq) * 100
  colnames(data) <- c("Level_of_Education", "Number_of_Patients", "Percentage")
  data <- data[order(data$Percentage), ]
  return(data)
}
```

```{r}
#calculating percentage data
Education_data <- calculate_percentage(Patient_Data, "Client_Education")

Home_Care_Edu <- Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  calculate_percentage("Client_Education")

In_patient_Edu <- Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  calculate_percentage("Client_Education")
```

```{r}
#creating a function to plot the bar graph
create_percentage_bar_graph <- function(data, y_col) {
  ggplot(data, aes(x = Level_of_Education, y = Percentage, fill = Level_of_Education)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(x = "Level of Education", y = y_col, title = paste("Percentage of", y_col, "by Education Level")) +
    scale_y_continuous(breaks = seq(0, 100, 10)) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))+
    geom_text(aes(label = sprintf("%.1f%%", Percentage)), hjust = -0.1)
}
```

```{r fig.width=13, fig.height=20}
plot_all_patients <- create_percentage_bar_graph(Education_data, "Patients")
plot_homecare <- create_percentage_bar_graph(Home_Care_Edu, "HomeCare Patients")
plot_inpatient <- create_percentage_bar_graph(In_patient_Edu, "InPatient Patients")

ggarrange(plot_all_patients, plot_homecare, plot_inpatient, 
          labels = c("All", "HomeCare", "InPatient"), 
          ncol = 1, nrow = 3)
```

# 6. Bar chart (When it was suggested that the patient requires a caregiver, what was your reaction)

```{r}
# Load necessary libraries
library(readxl)
library(dplyr)
library(stringr)
library(ggplot2)
library(ggpubr)

# Load the data
Patient_Data <- read_excel("Task2_data.xlsx")
as.data.frame(table(Patient_Data$Initial_Reaction)) 
```

```{r}
# Define the expected values for Initial_Reaction
reaction_levels <- c("Happy", "In different", "Didn’t Like it", "Was forced")

# Convert Initial_Reaction to factor with defined levels
Patient_Data$Initial_Reaction <- factor(Patient_Data$Initial_Reaction, levels = reaction_levels)
as.data.frame(table(Patient_Data$Initial_Reaction))  
```

```{r}
# Function to calculate percentages and format data
calculate_percentage <- function(data, column_name) {
  data <- as.data.frame(table(data[[column_name]]))
  data$Percentage = data$Freq / sum(data$Freq) * 100
  colnames(data) <- c("Reaction", "Frequency", "Percentage")
  data <- data[order(factor(data$Reaction, levels = reaction_levels)), ]
  return(data)
}

# Calculate percentage data for all patients
Initial_Reaction_Data <- calculate_percentage(Patient_Data, "Initial_Reaction")

# Calculate percentage data for Home Care patients
HomeCare_Initial_Reaction <- Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  calculate_percentage("Initial_Reaction")

# Calculate percentage data for Inpatient patients
Inpatient_Initial_Reaction <- Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  calculate_percentage("Initial_Reaction")
```

```{r}
# Function to create bar graph with percentage labels
create_percentage_bar_graph <- function(data, y_col) {
  ggplot(data, aes(x = Reaction, y = Percentage, fill = Reaction)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(x = "Initial Reaction", y = y_col, title = paste("Percentage of", y_col, "by Initial Reaction")) +
    scale_y_continuous(breaks = seq(0, 100, 10)) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))+
    geom_text(aes(label = sprintf("%.1f%%", Percentage)), hjust = -0.1)
}
```

```{r}
create_frequency_bar_graphs = function(data, x_col, y_col){
  y_lab = y_col
  x_lab =  x_col
  ttl = paste(y_lab, "by", x_lab)
  ggplot(data, aes_string(x = x_col, y = y_col, fill = x_col)) +
  geom_bar(stat = "identity") +
  coord_flip()+
  labs(x = x_lab, y = y_lab, title = ttl) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
}
```

```{r fig.width=13, fig.height=20}
# Create the plots
plot_all_patients <- create_frequency_bar_graphs(Initial_Reaction_Data, "Reaction", "Frequency")# create_percentage_bar_graph(Initial_Reaction_Data, "Patients")
plot_homecare <- create_frequency_bar_graphs(HomeCare_Initial_Reaction, "Reaction", "Frequency")# create_percentage_bar_graph(HomeCare_Initial_Reaction, "HomeCare Patients")
plot_inpatient <- create_frequency_bar_graphs(Inpatient_Initial_Reaction, "Reaction", "Frequency")# create_percentage_bar_graph(Inpatient_Initial_Reaction, "InPatient Patients")

# Arrange the plots using ggarrange
ggarrange(plot_all_patients, plot_homecare, plot_inpatient, 
          labels = c("All", "HomeCare", "Inpatient"), 
          ncol = 1, nrow = 3)
```

# Bar chart (%) Q47. what is your most preferred care arrangement

```{r}
Patient_Data = Patient_Data %>%
  mutate(Preferred_Care = str_replace_all(Preferred_Care, "Homecare provided by a Healthcare Provider", "Homecare Provided by a Healthcare Provider"),
         Preferred_Care = str_replace_all(Preferred_Care, "Homecare Provided by family member", "Homecare under a family member")
         )
```

```{r}
ordered_data <- Patient_Data %>%
  group_by(Preferred_Care) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  ungroup()
ordered_preferred_care <- ordered_data$Preferred_Care
ordered_preferred_care
```

```{r}
preferred_care_levels <- c("Homecare Provided by a Healthcare Provider", "Step down Facility", "Homecare under a family member", "Hospital Setting", "Any other facility that offer homecare")

calculate_percentage <- function(data, column_name) {
  data <- as.data.frame(table(data[[column_name]]))
  data$Percentage <- data$Freq / sum(data$Freq) * 100
  colnames(data) <- c("Preferred_Care", "Frequency", "Percentage")
  # Sort by Percentage
  data <- data[order(data$Percentage), ]
  # Ensure the Preferred_Care is ordered by predefined levels
  data$Preferred_Care <- factor(data$Preferred_Care, levels = preferred_care_levels)
  return(data)
}

# Sample Patient_Data for testing
Patient_Data <- data.frame(
  Preferred_Care = c("Homecare Provided by a Healthcare Provider", "Step down Facility", "Homecare under a family member", "Hospital Setting", "Any other facility that offer homecare", "Homecare Provided by a Healthcare Provider", "Hospital Setting"),
  Facility_Type = c("Home-care", "Home-care", "Inpatient", "Home-care", "Inpatient", "Inpatient", "Home-care")
)

# Calculate percentage data for all patients
Preferred_Care_Data <- calculate_percentage(Patient_Data, "Preferred_Care")

# Calculate percentage data for Home Care patients
HomeCare_Preferred_Care <- Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  calculate_percentage("Preferred_Care")

# Calculate percentage data for Inpatient patients
Inpatient_Preferred_Care <- Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  calculate_percentage("Preferred_Care")

# Display the results
Preferred_Care_Data
HomeCare_Preferred_Care
Inpatient_Preferred_Care

```

```{r fig.width=13, fig.height=20}
create_percentage_bar_graph <- function(data, y_col) {
  ggplot(data, aes(x = Preferred_Care, y = Percentage, fill = Preferred_Care)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(x = "Preferred Care", y = y_col, title = paste("Percentage of", y_col, "by Initial Reaction")) +
    scale_y_continuous(breaks = seq(0, 100, 10)) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))+
    geom_text(aes(label = sprintf("%.1f%%", Percentage)), hjust = -0.1)
}
# Create the plots
plot_all_patients <- create_percentage_bar_graph(Preferred_Care_Data, "Patients")
plot_homecare <- create_percentage_bar_graph(HomeCare_Preferred_Care, "Home-care Patients")
plot_inpatient <- create_percentage_bar_graph(Inpatient_Preferred_Care, "InPatient Patients")

# Arrange the plots using ggarrange
ggarrange(plot_all_patients, plot_homecare, plot_inpatient,
          labels = c("All", "B", "C"),
          ncol = 1, nrow = 3)
```

## 4. Bar chart for clients gross income

```{r}
# Load necessary libraries
library(readxl)
library(dplyr)
library(stringr)
library(ggplot2)
library(ggpubr)

# Load the data
Patient_Data <- read_excel("Task2_data.xlsx")
```

```{r}
# Clean and correct client income values
Patient_Data = Patient_Data %>%
  mutate(Client_Income = str_remove_all(Client_Income, "[[:space:]]"),
         Client_Income = str_replace_all(Client_Income, "5000,000", "500,000"),
         Client_Income = str_replace_all(Client_Income, "500000", "500,000"),
         Client_Income = str_replace_all(Client_Income, "200000", "200,000"),
         Client_Income = str_replace_all(Client_Income, "3000,000", "300,000"),
         Client_Income = str_replace_all(Client_Income, "101,000<150,000", "101,000-150,000"),
         Client_Income = str_replace_all(Client_Income, "100,000-", "101,000-"),
         Client_Income = str_replace_all(Client_Income, "200,000-", "201,000-"),
         Client_Income = str_replace_all(Client_Income, "300,000-", "301,000-"),
         Client_Income = str_replace_all(Client_Income, ">50,000", "<50,000"),
         Client_Income = str_replace_all(Client_Income, "51,000", "50,000"),
         Client_Income = str_replace_all(Client_Income, "10,000", "151,000"),
         Client_Income = ifelse(str_detect(Client_Income, ">500,000"), ">500,000", Client_Income),
         Client_Income = ifelse(str_detect(Client_Income, "<50,000"), "<50,000", Client_Income))
```

```{r}
as.data.frame(table(Patient_Data$Client_Income))
```

```{r}
# Define the correct order of income levels
income_levels <- c("<50,000", "50,000-100,000", "101,000-150,000", "151,000-200,000","201,000-300,000", "301,000-500,000", ">500,000")

# Convert Client_Income to factor with defined levels
Patient_Data$Client_Income <- factor(Patient_Data$Client_Income, levels = income_levels)
```

```{r}
# Function to calculate percentages and format data
calculate_percentage <- function(data, column_name) {
  data <- as.data.frame(table(data[[column_name]]))
  data$Percentage = data$Freq / sum(data$Freq) * 100
  colnames(data) <- c("Income", "Frequency", "Percentage")
  data <- data[order(factor(data$Income, levels = income_levels)), ]
  return(data)
}

# Calculate percentage data for all patients
Income_Data <- calculate_percentage(Patient_Data, "Client_Income")

# Calculate percentage data for Home Care patients
HomeCare_Income <- Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  calculate_percentage("Client_Income")

# Calculate percentage data for Inpatient patients
In_patient_Income <- Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  calculate_percentage("Client_Income")
```

```{r}
# Function to create bar graph with percentage labels
create_percentage_bar_graph <- function(data, y_col) {
  ggplot(data, aes(x = Income, y = Percentage, fill = Income)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(x = "Client Income", y = y_col, title = paste("Percentage of", y_col, "by Client Income")) +
    scale_y_continuous(breaks = seq(0, 100, 10)) +
    theme_minimal() +
    geom_text(aes(label = sprintf("%.1f%%", Percentage)), hjust = -0.1)
}
```

```{r fig.width=20, fig.height=13}
#  Create the plots
plot_all_patients <- create_percentage_bar_graph(Income_Data, "Patients")
plot_homecare <- create_percentage_bar_graph(HomeCare_Income, "HomeCare Patients")
plot_inpatient <- create_percentage_bar_graph(In_patient_Income, "InPatient Patients")

# Arrange the plots using ggarrange
ggarrange(plot_all_patients, plot_homecare, plot_inpatient, 
          labels = c("A", "B", "C"), 
          ncol = 1, nrow = 3)
```

# Correlate (Size of client household Q 32 and Clients gross income)

```{r}
Patient_Data = Patient_Data %>% 
  mutate(Client_Household = str_replace_all(Client_Household, "2to4", "2 to 4"))
```

```{r}
corr_data = data.frame(income = Patient_Data$Client_Income,household = Patient_Data$Client_Household)
table(corr_data)
```

```{r}
corr_data = corr_data %>% mutate(
  household = str_replace_all(household, ">7", "5<"),
  household = str_replace_all(household, "5 to 7", "5<"),
  household = str_replace_all(household, "1", "1 to 4"),
  household = str_replace_all(household, "2 to 4", "1 to 4"),
  income = str_replace_all(income, "<50,000", "0-100,000"),
  income = str_replace_all(income, "50,000-100,000", "0-100,000"),
)
corr_data = na.omit(corr_data)

table(corr_data)
```

The values to analyse do not include the row 3 and 4 because they have less than 5 on one entry

```{r}
table(corr_data)[-c(3, 4), ]

```

```{r}
corr_data = corr_data[corr_data$income %in% c(">500,000", " 0-100,000 ", "201,000-300,000 ", "301,000-500,000"),]
chisq.test(corr_data$income, corr_data$household)
```

income and household size have a statistically significant relationship

```{r}
library(vcd)
mosaicplot(table(corr_data), main = "Mosaic Plot of Client Income vs. Household Size", color = c("red", "blue"))
```

# Univariate regression for

```{r}
Patient_Data = Patient_Data %>%
  mutate(Alcohol = ifelse(Alcohol %in% c("Often", "More often", "Less extent"), "Others", Alcohol))
Patient_Data$Alcohol[1:10]

Patient_Data$Length_Separated = factor(Patient_Data$Length_Separated, levels = c("<6m", "1<3", "3<5", "More than 5 years"))
```

```{r}
HomeCare_data = Patient_Data %>% filter(Facility_Type == "Home-care")
InPatient_data = Patient_Data %>% filter(Facility_Type == "Inpatient")
```

```{r}
library(survival)
cox_regression = function(data, response){
  form = as.formula(paste("Surv(duration, event_observed) ~", response))
  rex_cox =  coxph(formula = form, data = data)
  return(summary(rex_cox))
}
```

## 1. Client Income

```{r}
print("cox regression for ALL PATIENTS Client Income")
cox_regression(Patient_Data, "Client_Income")
```

```{r}
print("cox regression for IN-PATIENTS Client Income")
cox_regression(InPatient_data, "Client_Income")
```

```{r}
print("cox regression for HOMECARE PATIENTS Client Income")
cox_regression(HomeCare_data, "Client_Income")
```

## 2. highest completed education level with categories 0-6

```{r}
cox_regression(Patient_Data, "Client_Education")
```

```{r}
cox_regression(InPatient_data, "Client_Education")
```

```{r}
cox_regression(HomeCare_data, "Client_Education")
```

> **Observation**\
>
> Secondary education is only significant in the case of `In-patient` care

## 3. How long have you been separated [0-4]

### Correct values

> Expected values
>
> -   last six months
>
> -   more than six(6) months but less than one(1) year
>
> -   1 year - below 3 years
>
> -   3 years - below 5 years
>
> -   more than 5 years

```{r}
as.data.frame(table(Patient_Data$Length_Separated))
```

```{r}
cox_regression(Patient_Data, "Length_Separated")
```

```{r}
cox_regression(HomeCare_data, "Length_Separated")
```

```{r}
cox_regression(InPatient_data, "Length_Separated")
```

## 4. Alcohol consumption [not at all 0, others 1]

```{r}
cox_regression(Patient_Data[!is.na(Patient_Data$Alcohol),], "Alcohol")
```

```{r}
cox_regression(HomeCare_data[!is.na(HomeCare_data$Alcohol),], "Alcohol")
```

```{r}
cox_regression(InPatient_data[!is.na(InPatient_data$Alcohol),], "Alcohol")
```

# Bar chart type of shift taken

Data is in the column `Shift_Taken`

The values expected are

> 24 hours (Day and Night)
>
> Day
>
> Night

```{r}
as.data.frame(table(Patient_Data$Shift_Taken))
```

### plot data

```{r}
shift_taken = as.data.frame(table(Patient_Data$Shift_Taken))
colnames(shift_taken) = c("shift", "frequency")  
shift_taken = shift_taken[shift_taken$shift %in% as.factor(c("Day", "Day and Night", "Night")),] 
shift_taken
```

### Plot

```{r}
create_frequency_bar_graphs(shift_taken,"shift", "frequency" )
```

### Shift by Homecare

```{r}
HomeCare_shift_taken = Patient_Data %>%   filter(Facility_Type == "Home-care") %>%   .$Shift_Taken %>%   table() %>%   as.data.frame() 
colnames(HomeCare_shift_taken) = c("shift_taken", "frequency")    
HomeCare_shift_taken 
```

### Shift by Income

```{r}
InPatient_shift_taken = Patient_Data %>%   filter(Facility_Type == "Inpatient") %>%   .$Shift_Taken %>%   table() %>%   as.data.frame() 
colnames(InPatient_shift_taken) = c("shift_taken", "frequency")    
InPatient_shift_taken 
create_frequency_bar_graphs(InPatient_shift_taken, "shift_taken", "frequency")
```

## Cross tabulate shift taken by GENDER and Marital status (two tables)

### cross tab function

```{r}
create_crosstab <- function(data, shift_col, res_col, facility_col) {
  
  # Helper function to process the data
  process_data <- function(data, shift_col, res_col, facility_filter = NULL) {
    if (!is.null(facility_filter)) {
      data <- data %>% filter(!!sym(facility_col) == facility_filter)
    }
    data <- data %>%
      filter(!is.na(!!sym(shift_col))) %>%
      select(!!sym(shift_col), !!sym(res_col))
    colnames(data) <- c("shift", "res")
    return(data)
  }
  
  # Process data for all patients
  shift_by <- process_data(data, shift_col, res_col)
  
  # Process data for home-care patients
  Homecare_shift_by <- process_data(data, shift_col, res_col, "Home-care")
  
  # Process data for inpatient patients
  InPatient_shift_by <- process_data(data, shift_col, res_col, "Inpatient")
  
  # Create cross-tabulations
  all_patient_crosstab <- as.data.frame(table(shift_by$res, shift_by$shift))
  colnames(all_patient_crosstab) <- c(res_col, "Shift", "Count")
  all_patient_crosstab <- all_patient_crosstab[order(all_patient_crosstab[[res_col]]), ]
  
  homecare_crosstab <- as.data.frame(table(Homecare_shift_by$res, Homecare_shift_by$shift))
  colnames(homecare_crosstab) <- c(res_col, "Shift", "Count")
  homecare_crosstab <- homecare_crosstab[order(homecare_crosstab[[res_col]]), ]
  
  inpatient_crosstab <- as.data.frame(table(InPatient_shift_by$res, InPatient_shift_by$shift))
  colnames(inpatient_crosstab) <- c(res_col, "Shift", "Count")
  inpatient_crosstab <- inpatient_crosstab[order(inpatient_crosstab[[res_col]]), ]
  
  return(list(
    all_patients = all_patient_crosstab,
    homecare_patients = homecare_crosstab,
    inpatient_patients = inpatient_crosstab
  ))
}
```

```{r}
gender_crosstab = create_crosstab(Patient_Data, "Shift_Taken", "Sex_Patient", "Facility_Type")
print("Cross tablulation of ALL patients")
gender_crosstab$all_patients
```

```{r}
print("Cross tablulation of in-patients")
gender_crosstab$inpatient_patients
```

```{r}
print("Cross tablulation of Home-care patients")
gender_crosstab$homecare_patients
```

### Marital status

```{r}
marital_crosstab = create_crosstab(Patient_Data, "Shift_Taken", "Marital", "Facility_Type")
```

```{r}
print("Cross tablulation of ALL patients")
marital_crosstab$all_patients
```

```{r}
print("Cross tablulation of in-patients")
marital_crosstab$inpatient_patients
```

```{r}
print("Cross tablulation of Home-care patients")
marital_crosstab$homecare_patients
```

## Q37. Cox regression for (How severe was the patient condition before engaging the caregiver)

The expected value for `Patient_Severity`

> -   Mild
>
> -   Severe
>
> -   Very Severe
>
> -   Not Applicable

```{r}
as.data.frame(table(Patient_Data$Patient_Severity))
```

```{r}
cox_regression(Patient_Data, "Patient_Severity")
```

## Also for qn40(current_condition...18) \*chisquared

```{r}
current_condition_crosstab = create_crosstab(Patient_Data, "Shift_Taken", "Current_Condition...18", "Facility_Type")
```

### All Patients

```{r}
current_condition_crosstab$all_patients %>% filter(Count > 5)
```

```{r}
library(tidyr)
qn40_all_pat = as.table(as.matrix(current_condition_crosstab$all_patients %>% filter(Count > 5) %>% pivot_wider(names_from = Shift, values_from = Count)))

chisq.test(table(qn40_all_pat))
```

### HomeCare

```{r}
current_condition_crosstab$homecare_patients%>% filter(Count > 5)
```

```{r}
as.table(as.matrix(current_condition_crosstab$homecare_patients%>% filter(Count > 5) %>% pivot_wider(names_from = Shift, values_from = Count)))
```

```{r}
qn40_home_pat = as.table(as.matrix(current_condition_crosstab$homecare_patients%>% filter(Count > 5) %>% pivot_wider(names_from = Shift, values_from = Count)))

chisq.test(table(qn40_home_pat))
```

### InPatient

```{r}
current_condition_crosstab$inpatient_patients%>% filter(Count > 5)
```

```{r}
as.table(as.matrix(current_condition_crosstab$inpatient_patients%>% filter(Count > 5) %>% pivot_wider(names_from = Shift, values_from = Count)))
```

> There is only one column so the chi-squared test will return `1` as there is nothing to compare with. also the In-Patient patients chose NA as a response for the current condition...18

```{r}
qn40_inpatient_pat = as.table(as.matrix(current_condition_crosstab$inpatient_patients%>% filter(Count > 5) %>% pivot_wider(names_from = Shift, values_from = Count)))

chisq.test(table(qn40_inpatient_pat))
```
