---
title: "Group 3 Task 2"
author: "Nzambuli Daniel & Waimiri Romeo"
date: "2024-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The data

```{r}
library(readxl)
library(dplyr)
library(stringr)
Patient_Data <- read_excel("Task2_data.xlsx")
head(Patient_Data)
```

# Bar chart(%) patient highest Completed Education Level

```{r}
Education_data = as.data.frame(table(Patient_Data$Client_Education))
Education_data

```

### Convert the data to percentages

```{r}
Education_data$percentage = Education_data$Freq/ sum(Education_data$Freq) * 100
Education_data
```

```{r}
colnames(Education_data) = c("Level_of_Edu", "No_Patients", "Percentage_Patients")
Education_data = Education_data[order(Education_data$Percentage_Patients),]
Education_data
```

### Plot the bargraph

```{r}
library(ggplot2)
ggplot(Education_data, aes(x = Level_of_Edu, y = Percentage_Patients, fill = Level_of_Edu)) +
  geom_bar(stat = "identity") +
  coord_flip()+
  labs(x = "Level of Education", y = "Percentage of Patients", title = "Percentage of Patients by Education Level") +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  theme_minimal()
```

### Plotting function

```{r}
create_frequency_bar_graphs = function(data, x_col, y_col){
  y_lab = y_col
  x_lab =  x_col
  ttl = paste(y_lab, "by", x_lab)
  ggplot(data, aes_string(x = x_col, y = y_col, fill = x_col)) +
  geom_bar(stat = "identity") +
  coord_flip()+
  labs(x = x_lab, y = y_lab, title = ttl) +
  theme_minimal()
}


create_percentage_bar_graphs = function(data, x_col, y_col){
  y_lab = y_col
  x_lab = x_col
  ttl = paste(y_lab, "by", x_lab)
  ggplot(data, aes_string(x = x_col, y = y_col, fill = x_col)) +
  geom_bar(stat = "identity") +
  coord_flip()+
  labs(x = x_lab, y = y_lab, title = ttl) +
  scale_y_continuous(breaks = seq(0, 100, 10)) + #, labels = scales::percent
  theme_minimal()
}
```

### Education by Home Care

```{r}
Home_Care_Edu = Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  .$Client_Education %>%
  table() %>%
  as.data.frame()

Home_Care_Edu$Percentage = Home_Care_Edu$Freq / sum(Home_Care_Edu$Freq) * 100
names(Home_Care_Edu) = c("Level_of_Edu", "No_Patients", "Perc_HomeCare_Patients")
Home_Care_Edu
```

```{r}
create_percentage_bar_graphs(Home_Care_Edu, "Level_of_Edu", "Perc_HomeCare_Patients")
```

### Education by in-patient

```{r}
In_patient_Edu = Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  .$Client_Education %>%
  table() %>%
  as.data.frame()

In_patient_Edu$Percentage = In_patient_Edu$Freq / sum(In_patient_Edu$Freq) * 100
names(In_patient_Edu) = c("Level_of_Edu", "No_Patients", "Perc_InPatient_Patients")
In_patient_Edu
```

```{r}
create_percentage_bar_graphs(In_patient_Edu, "Level_of_Edu", "Perc_InPatient_Patients")
```

# Univariate cox regression for patient

```{r}
library(survival)
cox_regression = function(data, response){
  form = as.formula(paste("Surv(duration, event_observed) ~", response))
  rex_cox =  coxph(formula = form, data = data)
  return(summary(rex_cox))
}
```

## 1. highest completed education level with categories 0-6

### Create columns of level of education

```{r}
as.character(Education_data$Level_of_Edu)
```

### Confirm if the values are in the original data set columns

```{r}
as.character(Education_data$Level_of_Edu) %in% colnames(Patient_Data)
```

### Make columns of 1 and 0

```{r}
create_indicator_column <- function(data, column_name, match_value, new_column_name) {
  data <- data %>%
    mutate(!!new_column_name := if_else(!!sym(column_name) == match_value, 1, 0))
  
  return(data)
}
```

### New columns

```{r}
columns_to_make = as.character(Education_data$Level_of_Edu)
for(column in columns_to_make){
  Patient_Data = create_indicator_column(Patient_Data, "Client_Education", column, column)
}
```

### Confirm

```{r}
Patient_Data$Secondary
```

## All Data set coxph

```{r}
names(Patient_Data) = sapply(names(Patient_Data), function(name) {
  if (name %in% columns_to_make) {
    
    gsub(" ", "_", name)
  } else {
    name
  }
})
```

```{r}
interested = gsub(" ", "_", columns_to_make)
for (column in interested){
  print(paste("cox regression for: ", column))
  cat("\n\n\n")
  print(cox_regression(Patient_Data, column))
  cat("\n\n\n")
}
```

## Home-stay coxph

```{r}
for (column in interested){
  filtered_data = Patient_Data %>% filter(Facility_Type != "Inpatient")
  print(paste("Home-stay cox regression for: ", column))
  cat("\n\n\n")
  print(cox_regression(filtered_data, column))
  cat("\n\n\n")
}
```

## In-patient coxph

```{r}
for (column in interested){
  filtered_data = Patient_Data %>% filter(Facility_Type == "Inpatient")
  print(paste("Inpatient cox regression for: ", column))
  cat("\n\n\n")
  print(cox_regression(filtered_data, column))
  cat("\n\n\n")
}
```

## 2. How long have you been separated [0-4]

in the column `Length_Separated`

```{r}
as.data.frame(table(Patient_Data$Length_Separated))
```

### Correct values

> Expected values
>
> -   last six months
>
> -   more than six(6) months but less than one(1) year
>
> -   1 year - below 3 years
>
> -   3 years - below 5 years
>
> -   more than 5 years

```{r}
Patient_Data = Patient_Data %>%
  mutate(Length_Separated = str_replace_all(Length_Separated, ">50,000", ">5"),
         Length_Separated = ifelse(Length_Separated == "NA", NA, Length_Separated)
         )
```

### confirm

```{r}
as.data.frame(table(Patient_Data$Length_Separated))
```

### Coxph data

```{r}
Patient_Data = Patient_Data %>%
  mutate(Length_Separated = ifelse(Length_Separated %in% c("<6m", ">5", "1<3", "3<5"), "Others", Length_Separated))
Patient_Data = create_indicator_column(Patient_Data, "Length_Separated", "Others", "Separated_more_than_4yrs")
```

### Confirm

```{r}
Patient_Data$Separated_more_than_4yrs
```

## All Data set coxph

```{r}
print(paste("cox regression for: ", "Separated_more_than_4yrs"))
cat("\n\n\n")
print(cox_regression(Patient_Data, "Separated_more_than_4yrs"))
```

## Home-stay coxph

```{r}
filtered_home_data = Patient_Data %>% filter(Facility_Type != "Inpatient")
print(paste("Home-stay cox regression for: ", "Separated_more_than_4yrs"))
cat("\n\n\n")
print(cox_regression(filtered_home_data, "Separated_more_than_4yrs"))
```

## In-patient coxph

```{r}
filtered_inpatient_data = Patient_Data %>% filter(Facility_Type == "Inpatient")
print(paste(" cox regression for: ", "Separated_more_than_4yrs"))
cat("\n\n\n")
print(cox_regression(filtered_inpatient_data, "Separated_more_than_4yrs"))
```

## 3. Alcohol consumption [not at all 0, others 1]

```{r}
as.data.frame(table(Patient_Data$Alcohol))
```

```{r}
sum(as.data.frame(table(Patient_Data$Alcohol))$Freq)
```

### coxph data

```{r}
Patient_Data = Patient_Data %>%
  mutate(Alcohol = ifelse(Alcohol %in% c("Often", "More often", "Less extent"), "Others", Alcohol))
Patient_Data$Alcohol[1:10]
```

```{r}
Patient_Data = create_indicator_column(Patient_Data, "Alcohol", "Others", "No_Alcohol")
```

### Confirm

```{r}
Patient_Data$No_Alcohol
```

## All Data set coxph

```{r}
print(paste("cox regression for: ", "No_Alcohol"))
cat("\n\n\n")
print(cox_regression(Patient_Data, "No_Alcohol"))
```

## Home-stay coxph

```{r}
filtered_home_data = Patient_Data %>% filter(Facility_Type != "Inpatient")
print(paste(" cox regression for: ", "No_Alcohol"))
cat("\n\n\n")
print(cox_regression(filtered_home_data, "No_Alcohol"))
```

## In-patient coxph

```{r}
filtered_inpatient_data = Patient_Data %>% filter(Facility_Type == "Inpatient")
print(paste(" cox regression for: ", "No_Alcohol"))
cat("\n\n\n")
print(cox_regression(filtered_inpatient_data, "No_Alcohol"))
```

## 4. Physical activities behavior [not at all 0, others 1]

```{r}
as.data.frame(table(Patient_Data$Physical_Activities))
```

```{r}
Patient_Data = Patient_Data %>%
  mutate(Physical_Activities = ifelse(Physical_Activities %in% c("Often", "More often", "Less extent"), "Others", Physical_Activities))
Patient_Data = create_indicator_column(Patient_Data, "Physical_Activities", "Others", "No_Physical_Activities")
```

## All Data set coxph

```{r}
print(paste("cox regression for: ", "No_Physical_Activities"))
cat("\n\n\n")
print(cox_regression(Patient_Data, "No_Physical_Activities"))
```

## Home-stay coxph

```{r}
filtered_home_data = Patient_Data %>% filter(Facility_Type != "Inpatient")

print(paste(" cox regression for: ", "No_Physical_Activities"))
cat("\n\n\n")
print(cox_regression(filtered_home_data, "No_Physical_Activities"))
```

## In-patient coxph

```{r}
filtered_inpatient_data = Patient_Data %>% filter(Facility_Type == "Inpatient")

print(paste(" cox regression for: ", "No_Physical_Activities"))
cat("\n\n\n")
print(cox_regression(filtered_inpatient_data, "No_Physical_Activities"))
```

## 5. Smoking [not at all 0, others 1]

```{r}
as.data.frame(table(Patient_Data$Smoking))
```

```{r}
Patient_Data = create_indicator_column(Patient_Data, "Smoking", "Yes", "No_Smoking")
```

## All Data set coxph

```{r}
print(paste("cox regression for: ", "No_Smoking"))
cat("\n\n\n")
print(cox_regression(Patient_Data, "No_Smoking"))
```

## Home-stay coxph

```{r}
filtered_home_data = Patient_Data %>% filter(Facility_Type != "Inpatient")

print(paste(" cox regression for: ", "No_Smoking"))
cat("\n\n\n")
print(cox_regression(filtered_home_data, "No_Smoking"))
```

## In-patient coxph

```{r}
filtered_inpatient_data = Patient_Data %>% filter(Facility_Type == "Inpatient")

print(paste(" cox regression for: ", "No_Smoking"))
cat("\n\n\n")
print(cox_regression(filtered_inpatient_data, "No_Smoking"))
```

# Correlate (Size of client household Q 32 and Clients gross income)

```{r}
as.data.frame(table(Patient_Data$Client_Household))
```

```{r}
Patient_Data = Patient_Data %>% 
  mutate(Client_Household = str_replace_all(Client_Household, "2to4", "2 to 4"))
```

```{r}
as.data.frame(table(Patient_Data$Client_Household))
```

```{r}
as.data.frame(table(Patient_Data$Client_Income))
```

### Correct values for client income

The values expected are

> \<50,000
>
> 51,000-100,000
>
> 101,000-150,000
>
> 151,000-200,000
>
> 201,000-300,000
>
> 300,000-500,000
>
> 500,000\<

```{r}
Patient_Data = Patient_Data %>%
  mutate(Client_Income = str_remove_all(Client_Income, "[[:space:]]"),
         Client_Income = str_replace_all(Client_Income, "5000,000", "500,000"),
         Client_Income = str_replace_all(Client_Income, "500000", "500,000"),
         Client_Income = str_replace_all(Client_Income, "200000", "200,000"),
         Client_Income = str_replace_all(Client_Income, "3000,000", "300,000"),
         Client_Income = str_replace_all(Client_Income, "101,000<150,000", "101,000-150,000"),
         Client_Income = str_replace_all(Client_Income, "100,000-", "101,000-"),
         Client_Income = str_replace_all(Client_Income, "200,000-", "201,000-"),
         Client_Income = str_replace_all(Client_Income, "300,000-", "301,000-"),
         Client_Income = str_replace_all(Client_Income, ">50,000", "<50,000"),
         Client_Income = str_replace_all(Client_Income, "51,000", "50,000"),
         Client_Income = str_replace_all(Client_Income, "10,000", "151,000"),
         Client_Income = ifelse(str_detect(Client_Income, ">500,000"), ">500,000", Client_Income),
         Client_Income = ifelse(str_detect(Client_Income, "<50,000"), "<50,000", Client_Income))
```

```{r}
as.data.frame(table(Patient_Data$Client_Income))
```

### The correlation

```{r}
corr_data = data.frame(income = Patient_Data$Client_Income,household = Patient_Data$Client_Household)
table(corr_data)
```

```{r}
corr_data = corr_data %>% mutate(
  household = str_replace_all(household, ">7", "5<"),
  household = str_replace_all(household, "5 to 7", "5<"),
  household = str_replace_all(household, "1", "1 to 4"),
  household = str_replace_all(household, "2 to 4", "1 to 4"),
  income = str_replace_all(income, "<50,000", "0-100,000"),
  income = str_replace_all(income, "50,000-100,000", "0-100,000"),
)
corr_data = na.omit(corr_data)

table(corr_data)
```

```{r}
# correlation = cor(Patient_Data$Client_Household, Patient_Data$Client_Income, use = "complete.obs") 
# correlation

chisq.test(corr_data$income, corr_data$household)
```

income and household size have a statistically significant relationship

```{r fig.width= 13}
library(vcd)
mosaicplot(table(corr_data), main = "Mosaic Plot of Client Income vs. Household Size", color = TRUE)

```

# Bar chart for clients gross income

Data is in the column `Client_Income`

The values expected are

> \<50,000
>
> 51,000-100,000
>
> 101,000-150,000
>
> 151,000-200,000
>
> 201,000-300,000
>
> 300,000-500,000
>
> 500,000\<

```{r}
Income_Data = Patient_Data$Client_Income
```

### Check the data is clean

```{r}
as.data.frame(table(Patient_Data$Client_Income))
```

### Relabel values

### Confirm

```{r}
as.data.frame(table(Patient_Data$Client_Income))
```

### Create plot data

```{r}
customer_income = as.data.frame(table(Patient_Data$Client_Income))
colnames(customer_income) = c("Income", "Frequency")
```

```{r}
create_frequency_bar_graphs(customer_income, "Income", "Frequency")
```

### Income by in-patient

```{r}
In_patient_Income = Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  .$Client_Income %>%
  table() %>%
  as.data.frame()


names(In_patient_Income) = c("Client_Income", "InPatient_Freq")
In_patient_Income
```

```{r}
create_frequency_bar_graphs(In_patient_Income, "Client_Income", "InPatient_Freq")
```

### Income by home-care

```{r}
HomeCare_Income = Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  .$Client_Income %>%
  table() %>%
  as.data.frame()


names(HomeCare_Income) = c("Client_Income", "HomeCare_Freq")
HomeCare_Income
create_frequency_bar_graphs(HomeCare_Income, "Client_Income", "HomeCare_Freq")
```

# Univariate regression for

## 1. Client Income

```{r}
Patient_Data <- Patient_Data %>%
  mutate(
    income_gt_500000 = as.integer(Client_Income == ">500,000"),
    income_301000_500000 = as.integer(Client_Income == "301,000-500,000"),
    income_101000_150000 = as.integer(Client_Income == "101,000-150,000"),
    income_150000_200000 = as.integer(Client_Income == "150,000-200,000"),
    income_lt_50000 = as.integer(Client_Income == "<50,000"),
    income_201000_300000 = as.integer(Client_Income == "201,000-300,000"),
    income_151000_200000 = as.integer(Client_Income == "151,000-200,000")
  )
```

### Cox regression

```{r}
income_cols = c("income_gt_500000", "income_301000_500000", "income_101000_150000", "income_150000_200000", "income_lt_50000", "income_201000_300000", "income_151000_200000")

for(column in income_cols){
  # filtered_data = Patient_Data %>% filter(Facility_Type == "Inpatient")
  print(paste("Inpatient cox regression for: ", column))
  cat("\n\n\n")
  print(cox_regression(Patient_Data, column))
  cat("\n\n\n")
}
```

### Home-care

```{r}
for(column in income_cols){
  filtered_data = Patient_Data %>% filter(Facility_Type != "Inpatient")
  print(paste("Home-care cox regression for: ", column))
  cat("\n\n\n")
  print(cox_regression(filtered_data, column))
  cat("\n\n\n")
}
```

### In-patient

```{r}
for(column in income_cols){
  filtered_data = Patient_Data %>% filter(Facility_Type == "Inpatient")
  print(paste("Inpatient cox regression for: ", column))
  cat("\n\n\n")
  print(cox_regression(filtered_data, column))
  cat("\n\n\n")
}
```

## 2. Q37. (How severe was the patient condition before engaging the caregiver)

# Bar chart (When it was suggested that the patient requires a caregiver, what was your reaction)

Data is in the column `Initial_Reaction`

The values expected are

> Happy
>
> In different
>
> Didn't like it at all
>
> Was Forced

```{r}
as.data.frame(table(Patient_Data$Initial_Reaction))
```

### Plot data

```{r}
initial_reaction = as.data.frame(table(Patient_Data$Initial_Reaction))
colnames(initial_reaction) = c("Client_Reaction", "Frequency")
initial_reaction
```

### Plot

```{r}
create_frequency_bar_graphs(initial_reaction,"Client_Reaction", "Frequency" )
```

### Initial reaction by homecare

```{r}
HomeCare_Initial_Reaction = Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  .$Initial_Reaction %>%
  table() %>%
  as.data.frame()


names(HomeCare_Initial_Reaction) = c("Client_Initial_Reaction", "HomeCare_Freq")
HomeCare_Initial_Reaction
create_frequency_bar_graphs(HomeCare_Initial_Reaction, "Client_Initial_Reaction", "HomeCare_Freq")
```

### Initial reaction by inpatient

```{r}
Inpatient_Initial_Reaction = Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  .$Initial_Reaction %>%
  table() %>%
  as.data.frame()


names(Inpatient_Initial_Reaction) = c("Client_Initial_Reaction", "Inpatient_Freq")
Inpatient_Initial_Reaction
create_frequency_bar_graphs(Inpatient_Initial_Reaction, "Client_Initial_Reaction", "Inpatient_Freq")
```

# Bar chart (%) Q47. what is your most preferred care arrangement

Data is in the column `Preferred_care`

The values expected are

> Homecare under a family member/relative
>
> Homecare provided by a healthcare provider
>
> Hospital setting
>
> Step down facility like Care 360
>
> Other

```{r}
as.data.frame(table(Patient_Data$Preferred_Care))
```

combine

> Homecare Provided by a Healthcare Provider -\> Homecare provided by a Healthcare Provider
>
> Homecare Provided by family member -\> Homecare under a family member

```{r}
Patient_Data = Patient_Data %>%
  mutate(Preferred_Care = str_replace_all(Preferred_Care, "Homecare provided by a Healthcare Provider", "Homecare Provided by a Healthcare Provider"),
         Preferred_Care = str_replace_all(Preferred_Care, "Homecare Provided by family member", "Homecare under a family member")
         )
```

### Confirm

```{r}
as.data.frame(table(Patient_Data$Preferred_Care))
```

### plot data

```{r}
preferred_care = as.data.frame(table(Patient_Data$Preferred_Care))
colnames(preferred_care) = c("preferred_care", "frequency")
preferred_care$percentage = preferred_care$frequency/ sum(preferred_care$frequency) * 100
preferred_care
```

### Plot

```{r fig.width=13}
create_percentage_bar_graphs(preferred_care, "preferred_care", "percentage")
```

### Preferred care for inpatient

```{r}
Inpatient_Preferred_Care = Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  .$Preferred_Care %>%
  table() %>%
  as.data.frame()
colnames(Inpatient_Preferred_Care) = c("preferred_care", "frequency")
Inpatient_Preferred_Care$Inpatient_percentage = Inpatient_Preferred_Care$frequency/ sum(Inpatient_Preferred_Care$frequency) * 100


Inpatient_Preferred_Care
create_percentage_bar_graphs(Inpatient_Preferred_Care, "preferred_care", "Inpatient_percentage")
```

### preferred care by Homecare

```{r fig.width=13}
HomeCare_Preferred_Care = Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  .$Preferred_Care %>%
  table() %>%
  as.data.frame()
colnames(HomeCare_Preferred_Care) = c("preferred_care", "frequency")
HomeCare_Preferred_Care$HomeCare_percentage = HomeCare_Preferred_Care$frequency/ sum(HomeCare_Preferred_Care$frequency) * 100


HomeCare_Preferred_Care
create_percentage_bar_graphs(HomeCare_Preferred_Care, "preferred_care", "HomeCare_percentage")
```

# Bar chart type of shift taken

Data is in the column `Shift_Taken`

The values expected are

> 24 hours (Day and Night)
>
> Day
>
> Night

```{r}
as.data.frame(table(Patient_Data$Shift_Taken))
```

### plot data

```{r}
shift_taken = as.data.frame(table(Patient_Data$Shift_Taken))
colnames(shift_taken) = c("shift", "frequency")

shift_taken = shift_taken[shift_taken$shift %in% as.factor(c("Day", "Day and Night", "Night")),]
shift_taken
```

### Plot

```{r}
create_frequency_bar_graphs(shift_taken,"shift", "frequency" )
```

### Shift by Homecare

```{r}
HomeCare_shift_taken = Patient_Data %>%
  filter(Facility_Type == "Home-care") %>%
  .$Shift_Taken %>%
  table() %>%
  as.data.frame()
colnames(HomeCare_shift_taken) = c("shift_taken", "frequency")



HomeCare_shift_taken

```

### Shift by Income

```{r}
InPatient_shift_taken = Patient_Data %>%
  filter(Facility_Type == "Inpatient") %>%
  .$Shift_Taken %>%
  table() %>%
  as.data.frame()
colnames(InPatient_shift_taken) = c("shift_taken", "frequency")



InPatient_shift_taken
create_frequency_bar_graphs(InPatient_shift_taken, "shift_taken", "frequency")
```
