---
title: "Survival analysis with veteran data"
author: "Nzambuli Daniel"
date: "2024-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r}
library(survival)
library(ranger)
library(ggplot2)
library(dplyr)
library(ggfortify)

```

```{r}
data = veteran
head(data)
```

> **Describing the data**
>
> |               |                                           |
> |:--------------|:------------------------------------------|
> | **trt:**      | 1=standard 2=test                         |
> | **celltype:** | 1=squamous, 2=smallcell, 3=adeno, 4=large |
> | **time:**     | survival time                             |
> | **status:**   | censoring status                          |
> | **karno:**    | Karnofsky performance score (100=good)    |
> | **diagtime:** | months from diagnosis to randomisation    |
> | **age:**      | in years                                  |
> | **prior:**    | prior therapy 0=no, 10=yes                |

## Kaplan Meier Analysis

```{r}
km = with(data, Surv(time, status))
head(km, 80)
```

> **Pluses –** indicate that censoring has taken place

## Km - fit

```{r}
km_fit = survfit(Surv(time, status)~1, data = data) # for all the data set, link = 1
summary(km_fit)

```

### select only specific colums

> select the values at point `1, 30, 60, 90, 180, 270, ...`

```{r}
summary(km_fit, times = c(1, 30, 60, 90 * (1:10)))
```

## Plot

```{r}
plot(km_fit, xlab = "Days", main = "Kaplan Meier Plot")
```

### Using autoplot

```{r}
autoplot(km_fit)
```

## Treatment

```{r}
km_fit_trt = survfit(Surv(time, status)~trt, data = data) # for all the data set, link = 1
autoplot(km_fit_trt)
```

## Trying as a factor

```{r}
data_2  = mutate(data, trt = as.factor(ifelse(trt == 1,"standard", "test")))

km_fit_trt = survfit(Surv(time, status)~trt, data = data_2) # for all the data set, link = 1
autoplot(km_fit_trt)
```

## Survival by age

```{r}
vet = mutate(data, AG = ifelse((age < 60), "LT60", "OV60"),
             AG = factor(AG),
             trt = factor(trt, labels = c("standard", "test")),
             prior = factor(prior, labels = c("No", "Yes")))

km_AG_fit = survfit(Surv(time, status)~AG, data = vet)
autoplot(km_AG_fit)
```

## GGsurvfit fit the survival model

```{r}
library(ggsurvfit)
```

# Using df_colon data

|           |                                                                                   |
|:-----------------|:-----------------------------------------------------|
| id:       | id                                                                                |
| study:    | 1 for all patients                                                                |
| rx:       | Treatment - Obs(ervation), Lev(amisole), Lev(amisole)+5-FU                        |
| sex:      | 1=male                                                                            |
| age:      | in years                                                                          |
| obstruct: | obstruction of colon by tumour                                                    |
| perfor:   | perforation of colon                                                              |
| adhere:   | adherence to nearby organs                                                        |
| nodes:    | number of lymph nodes with detectable cancer                                      |
| time:     | days until event or censoring                                                     |
| status:   | censoring status                                                                  |
| differ:   | differentiation of tumour (1=well, 2=moderate, 3=poor)                            |
| extent:   | Extent of local spread (1=submucosa, 2=muscle, 3=serosa, 4=contiguous structures) |
| surg:     | time from surgery to registration (0=short, 1=long)                               |
| node4:    | more than 4 positive lymph nodes                                                  |
| etype:    | event type: 1=recurrence,2=death                                                  |

```{r}
colon_data = df_colon

p = survfit(Surv(time, status)~surg, data = colon_data)|>
  ggsurvfit(linewidth = 1)+
  add_confidence_interval()+
  add_risktable()+
  add_quantile(y_value = 0.6, color = "gray50", linewidth = 0.75)+
  scale_ggsurvfit()

p+
  # limit plot to show 8 years or less
  coord_cartesian(xlim = c(0, 8))+
  # update figure labels/titles
  labs(
    y = "percentage survival",
    title = "recurrence by time from surgery to randomization"
  )
```

## New model

```{r}
library(tibble)
library(knitr)
library(ezfun)
library(lubridate)
library(gtsummary)
library(tidycmprsk)
library(condSURV)
```

```{r}
lung = lung %>% mutate(status = recode(status, `1` = 0, `2` = 1))
head(lung)
```

# Sx_date data

```{r}
date_ex = tibble(sx_date = c("2007-06-22", "2004-02-13", "2010-10-27"),
                 last_fup_date = c("2017-04-15", "2018-07-04", "2016-10-31"))
date_ex
```

```{r}
date_ex = date_ex %>% mutate(
  sx_date = ymd(sx_date),
  last_fup_date = ymd(last_fup_date)
)
date_ex
```

```{r}

```
